---
title: "Assignment 1"
author: "Holly Finertie (HF2379)"
date: "1/17/2020"
output: github_document
---

```{r}
library(tidyverse)
library(ggplot2)
```


## Problem 1: 

Please find tables below that provide summaries (min, mean, median, IQR, max) of the quantitative features of the dataset. 

```{r}
data = read_csv("./data/data.csv") %>% 
  janitor::clean_names()

summary(data) 
```

## Problem 2:   

Using code below, the continuous BMI variable was transformed into a categorical variable (bmi_cat) with the following BMI categories: 

* Severely underweight: BMI < 16.5kg/m^2 
* Underweight: 16.5 <= BMI <= 18.5 kg/m^2 
* Normal weight: 18.5 <= BMI <=24.9 kg/m^2 
* Overweight: 25 <= BMI <= 29.9 kg/m^2 
* Obesity class I: 30 <= BMI <= 34.9 kg/m^2 
* Obesity class II: 35 <= BMI <= 39.9 kg/m^2 
* Obesity class III: BMI >= 40 kg/m^2

```{r}
data_bmi = data %>% 
  mutate(bmi_cat = as.factor(case_when(
      bmi <= 16.4 ~ "Severely underweight", 
      bmi >= 16.5 & bmi <= 18.4 ~ "Underweight",
      bmi >=18.5 & bmi <= 24.9 ~ "Normal weight",
      bmi >= 25 & bmi <= 29.9 ~ "Overweight", 
      bmi >= 30 & bmi <= 34.9 ~ "Obesity 1", 
      bmi >= 35 & bmi <= 39.9 ~ "Obesity 2", 
      bmi >= 40 ~ "Obesity 3"
    )), 
    bmi_cat = fct_relevel(
      bmi_cat, 
        str_c(c("Underweight", "Normal weight", 
               "Overweight", "Obesity 1", "Obesity 2")))
    )
    
  
```


## Problem 3: 


```{r}
data_final = data_bmi %>% 
  mutate(
    Arm = recode(classification, 
      `1` = "control", 
      `2` = "case"), 
    outcome = recode(Arm, 
      "control" = 0, 
      "case" = 1))
    

plot = data_final %>% 
  ggplot(aes(x = bmi_cat, fill = Arm)) +
  geom_bar(stat = "count") + 
  xlab("BMI Category") +
  ylab("Count of Breast CAncer Cases and Controls") +
  labs(
    title = "Proportion of Breast Cancer Cases and Controls by BMI Category"
  )

plot
```


## Problem 4: 


```{r}
logit_reg = 
  glm(outcome ~ glucose + homa + leptin + bmi + age, 
      family = binomial(link = "logit"), data = data_final) %>% 
  broom::tidy() %>% 
  mutate(
      "Lower Limit" = estimate - (std.error*1.96), 
      "Upper Limit" = estimate + (std.error*1.96)
  ) %>% filter(term == "homa") %>% 
  select(term, estimate, "Lower Limit", "Upper Limit") %>% 
  knitr::kable()


logit_reg
```

As seen in the table above, the beta estimate associated with a 1-unit change in HOMA is 0.274 with 95% CI(-0.063, 0.611). 


## Problem 5: 
```{r}
linear_reg = 
  lm(insulin ~ bmi + age + glucose, data = data_final) %>% 
  broom::tidy() %>% 
  mutate(
      "Lower Limit" = estimate - (std.error*1.96), 
      "Upper Limit" = estimate + (std.error*1.96)
  ) %>% filter(term == "age") %>% 
  select(term, estimate, "Lower Limit", "Upper Limit") %>% 
  knitr::kable()

linear_reg
```

As seen in the table above, the beta estimate associated with a 1-unit change in age is -0.054 with 95% CI(-0.156, 0.048). 


